Optimizing Pypy's Garbage Collector for Copy On Write Performance given a forky Web-Server Workload

### 2018-03-01 paper/main.tex
Added intro.  Fixed formatting

  \section{Introduction}
  Advances in interpretors for high level languages have enabled the use of high level languages in performance critical applications.  Some of the most visited websites on the Internet as of 2018 are written in high level languages such as Python, Ruby, and PHP.  A common problem with these deployments is memory usage: garbage collected languages tend to use more memory than their manually-memory-managed counterparts.  Most research into garbage collection focuses on the process in isolation.  This paper, however, examines garbage collection in a group of processes which share memory in a Copy-On-Write fashion.

  \begin{enumerate}
	\item We are looking at web servers.  Web servers fork a lot.
	\item Forked processes share CoW memory... garbage collection is messy
	\item We change garbage collector to optimize CoW by making mature pointers all indirect so mature objects don't get touched during minor collection
  \end{enumerate}

  \section{Background}

@@

  \section{Design}

@@

  \section{Implementation}
  \begin{enumerate}
	\item Give a summary of which structures and functions actually changed
	\item Where did I store this information?  How is it accessed in code?  
	\item maybe some pseudocode snippets
  \end{enumerate}

  \section{Methodologies}
  \begin{enumerate}
	\item Ngnx + uWSGI + Django + pypy in multi-process mode
	\item Use thebluealliance.com as a test load becaues it caches a lot of data and does some non-trivial computation
	\item Generate a bunch of fake traffic using httperf
	\item Monitor running code
	  \begin{enumerate}
        \item How long did it take to handle 10,000 requests
        \item What was the average time-to-last-byte?
		\item Total memory of all processes
		\item Average memory of each process
		\item shared memory per process
		\item num page faults
		\item cache miss rate?
		\item pages served per second (did it go up or down?)
		\item How many times did major and minor garbage collection happen?
		\item How many times did this inter-generation optimization actually occur?
	  \end{enumerate}
  \end{enumerate}

  \section{Analysis}
  \begin{enumerate}
	\item Did we actually save any memory?  How much on average?  Were there any patterns?   
	\item How often did old objects actually refer to young objects?  
	\item How often did minor collection happen?  Major collection?
	\item How much CPU overhead was there in following that extra pointer every time?  Can I even measure this?
	\item How many pages did we serve per second?  Did it go up or down?
  \end{enumerate}

  \section{Future Work}
  \begin{enumerate}
	\item What other changes could I make to improve CoW performance?  Lemme think.
  \end{enumerate}

  \section{Related work}

@@

  The optimization described in this paper is different in that we focus on preserving old objects when objects are copied from eden into the older generation.

  \subsection{Copy-On-Write Friendly Garbage Collection in CPython at Instagram}
  A 2017 article from Instagram's Engineering series describes changes to CPython's garbage collector made at instagram aimed at improving memory performance in their web-tier servers.  The experimentation section of this paper is heavily inspired by the test setup used at Instagram.

  The optimization implemented at instagram was to have the programmer manually mark certain objects as "GC-Frozen".  Objects that are GC-Frozen are not touched by the garbage collector and so the data in and around them is not touched.  Because Instagram's specific use case involves a lot of static objects that are shared between processes, this worked effectively.

>>>>

\section{Introduction}
Advances in interpretors for high level languages have enabled the use of high level languages in performance critical applications.  Some of the most visited websites on the Internet as of 2018 are written in high level languages such as Python, Ruby, and PHP.  A common problem with these deployments is memory usage: garbage collected languages tend to use more memory than their manually-memory-managed counterparts. In order for these popular websites to continue to scale up, the tools on which they are developed need to get more efficient.  Webservers need to use less memory.

Most research into garbage collection focuses on the process in isolation.  This paper, however, examines garbage collection in a group of processes which share memory in a Copy-On-Write fashion.  In a webserver environment where worker processes are often forked repeatedly during operation, this situation is very common.  Workers forked from the master process initially share some memory with the master and so the memory usage of the system as a whole is directly related to how much memory can be shared between processes.  This is the type of environment that many of today's largest websites like Instagram\cite{dismissing_garbage} operate with and so we believe that improvements made in this area can be used to improve performance in the field as a whole.

To accomplish this goal, this paper examines an optimization in PyPy's Generational Garbage Collector which is designed to reduce the number of shared pages that get dirtied during the copying phase of a minor collection.  Section \ref{sec:background} explains various topics relating to the optimization and builds up to a demonstration intended to create some intuition into how this optimziation is designed to work.  Section \ref{sec:design} explains the nitty gritty details of the optimization and builds intuition as to which data structures need to be changed.  Section \ref{sec:implementation} summarizes the source changes made to PyPy's garbage collector in order to implement this.  Section \ref{sec:method} talks about how these changes were tested and section \ref{sec:analysis} examines the results of these tests and explains them.  Finally, section \ref{sec:relatedwork} talks about other optimizations that have been made to Garbage Collectors to optimize for Copy-On-write behavior and explains how they differ from this research.

\section{Background}\label{sec:background}

@@

\section{Design}\label{sec:design}

@@

\section{Implementation}\label{sec:implementation}
\begin{enumerate}
    \item Give a summary of which structures and functions actually changed
    \item Where did I store this information?  How is it accessed in code?  
    \item maybe some pseudocode snippets
\end{enumerate}

\section{Methodologies}\label{sec:method}
\begin{enumerate}
    \item Ngnx + uWSGI + Django + pypy in multi-process mode
    \item Use thebluealliance.com as a test load because it caches a lot of data and does some non-trivial computation
    \item Generate a bunch of fake traffic using httperf
    \item Monitor running code
        \begin{enumerate}
            \item How long did it take to handle 10,000 requests
            \item What was the average time-to-last-byte?
            \item Total memory of all processes
            \item Average memory of each process
            \item shared memory per process
            \item num page faults
            \item cache miss rate?
            \item pages served per second (did it go up or down?)
            \item How many times did major and minor garbage collection happen?
            \item How many times did this inter-generation optimization actually occur?
        \end{enumerate}
\end{enumerate}

\section{Analysis}\label{sec:analysis}
\begin{enumerate}
    \item Did we actually save any memory?  How much on average?  Were there any patterns?   
    \item How often did old objects actually refer to young objects?  
    \item How often did minor collection happen?  Major collection?
    \item How much CPU overhead was there in following that extra pointer every time?  Can I even measure this?
    \item How many pages did we serve per second?  Did it go up or down?
\end{enumerate}

\section{Future Work}\label{sec:futurework}
\begin{enumerate}
    \item What other changes could I make to improve CoW performance?  Lemme think.
\end{enumerate}

\section{Related work}\label{sec:relatedwork}

@@

The optimization described in this paper is different in that we focus on preserving old objects when objects are copied from Eden into the older generation.

\subsection{Copy-On-Write Friendly Garbage Collection in CPython at Instagram}
A 2017 article from Instagram's Engineering series describes changes to CPython's garbage collector made at Instagram aimed at improving memory performance in their web-tier servers.  The experimentation section of this paper is heavily inspired by the test setup used at Instagram.

The optimization implemented at Instagram was to have the programmer manually mark certain objects as "GC-Frozen".  Objects that are GC-Frozen are not touched by the garbage collector and so the data in and around them is not touched.  Because Instagram's specific use case involves a lot of static objects that are shared between processes, this worked effectively.
